---
title: "MS&E 226 Project"
author: "Peter Karnchanapimonkul (SUNET: phatk), Tyler Coleman (SUNET: colemant)"
always_allow_html: true
output:
  pdf_document: default

---
```{r include=FALSE}
# install.packages("glmnet")
library(plotmo) # for plot_glmnet
library(glmnet)
library(cvTools)
library(tidyverse)
library(tibble)
library(plotly)
library(ggplot2)
library(rlang)
library(mosaicData)
library(caret)
options(scipen = 999, digits = 5)
library(stringr)
library(e1071)
library(ada)
library(class)
library(parallelSVM)
```

# IMPORT DATA AND SPLIT INTO TRAIN AND HOLDOUT

```{r}
setwd("/Users/tylercoleman/Desktop/")
NFL_DATA <- read.csv(file = "train.csv", header = TRUE, sep=",")
NFL_DATA_Run_Observations <- NFL_DATA[(NFL_DATA$NflIdRusher == NFL_DATA$NflId)   , ]

# split 80:20 for training:test
set.seed(123)
training_data = sample(nrow(NFL_DATA_Run_Observations), size = nrow(NFL_DATA_Run_Observations) * 0.8) 
NFL_DATA_Train = NFL_DATA_Run_Observations[training_data, ] 
NFL_DATA_Holdout = NFL_DATA_Run_Observations[-training_data, ] # holdout is remaining indices

#View(NFL_DATA_Train) 
```

# DATA MANIPULATION AND CLEANING
Adding the covariates that we want to include, and modififying the dataframe 

```{r}
NFL_DATA_TRAIN_Modified <- NFL_DATA_Train

# Function to take the difference in time from the dataframe
timeDifference <- function(time) {
  num <- gsub("[:]", "" , str_sub(time, 12, 19), perl=TRUE)
  hr <- ifelse(str_sub(num, 1, 2) == "00", 24, as.numeric(str_sub(num, 1, 2)))
  min <- as.numeric(str_sub(num, 3, 4))
  sec <- as.numeric(str_sub(num, 5, 6))
  newTime <- 3600*hr + 60 * min + sec
  return(newTime)
}
# Add Time_Difference between the snap and the handoff
NFL_DATA_TRAIN_Modified$TimeDifference <- 
  timeDifference(NFL_DATA_TRAIN_Modified$TimeHandoff) - timeDifference(NFL_DATA_TRAIN_Modified$TimeSnap)

# Add the Difference in Score by home score - visitor score
# Difference in Score (Pair with which team is winning (HomeScore-AwayScore))
NFL_DATA_TRAIN_Modified$HomeScoreAdvantage <- 
  NFL_DATA_TRAIN_Modified$HomeScoreBeforePlay - NFL_DATA_TRAIN_Modified$VisitorScoreBeforePlay

# Add the age of the running player

# Change the birth dates to strings
NFL_DATA_TRAIN_Modified$PlayerBirthDate = as.character(NFL_DATA_TRAIN_Modified$PlayerBirthDate)
# Grab the Year for each of the running player
Birth_Year = str_sub(NFL_DATA_TRAIN_Modified$PlayerBirthDate, 7, 11)
# Grab Month of each running player
Birth_Month = str_sub(NFL_DATA_TRAIN_Modified$PlayerBirthDate, 1, 2)
# If Born in July (07) Have lived 5/12 of a year. ie (12 - (Birth_Month)) / 12
How_Much_Of_Year_Lived = (12 - as.numeric(Birth_Month)) / 12
Years_Lived = NFL_DATA_TRAIN_Modified$Season - as.numeric(Birth_Year)
Total_Years_Lived = Years_Lived + How_Much_Of_Year_Lived
NFL_DATA_TRAIN_Modified$PlayerAge = Total_Years_Lived

# Change HEIGHT to inches and continuous
Feet = as.numeric(str_sub(NFL_DATA_TRAIN_Modified$PlayerHeight, 1, 1)) 
Inches = as.numeric(str_sub(NFL_DATA_TRAIN_Modified$PlayerHeight, 3, 4)) 
Heights = (Feet * 12) + Inches
NFL_DATA_TRAIN_Modified$PlayerHeight = Heights

# Changes GAMECLOCK to Seconds. 

NFL_DATA_TRAIN_Modified$GameClock = as.numeric(NFL_DATA_TRAIN_Modified$GameClock)

# FACTORING VARIABLES INTO CATEGORICAL

# Factor OFFENSE FORMATION 
NFL_DATA_TRAIN_Modified$OffenseFormation = factor(NFL_DATA_TRAIN_Modified$OffenseFormation) 

# DEFENDERS IN BOX (Need Categorical and Ordinal)
NFL_DATA_TRAIN_Modified$DefendersInTheBox = factor(NFL_DATA_TRAIN_Modified$DefendersInTheBox)  


# REFACTOR DEFENDER IN BOX TO INCLUDE ORDINALITY
# Leaving them unordered was for graphs above, to look good
NFL_DATA_TRAIN_Modified$DefendersInTheBox = factor(NFL_DATA_TRAIN_Modified$DefendersInTheBox, order = TRUE, levels= c(2,3,4,5,6,7,8,9,10,11)) 
```

# SELECTION OF COVARIATES FOR ANALYSIS
```{r}
# Drop columns that are collinear, or we think are not critical to our model
NFL_DATA_TRAIN_Filtered = select(NFL_DATA_TRAIN_Modified, 
                                -GameId, -PlayId, -Team, -S, -A, -Dis,
                                -Orientation, -Dir, -DisplayName, -JerseyNumber,
                                -YardLine, -FieldPosition, -HomeScoreBeforePlay,
                                -VisitorScoreBeforePlay, -NflId, -TimeHandoff,
                                -TimeSnap, -PlayerBirthDate, -PlayerCollegeName, 
                                -Location,-WindSpeed, -WindDirection, -StadiumType,
                                -Turf,-GameWeather,-NflIdRusher, -Stadium) 
# Turf and stadium type all captured in stadium
# View(NFL_DATA_TRAIN_Filtered) # drop game weather, captured in Week, and Stadium. Also too many missing values
```

# NOW THAT HAVE SELECTED COVARIATES. MORE DATA CLEANING, FACTORING, ETC...
```{r}
# Need to count how many NA / Empty cells there are for each column
# summary(NFL_DATA_TRAIN_Filtered)  # changed empty to NA when reading in file
# GameWeather, Temperature, Humidy all have missing or NA data
# sum(is.na(NFL_DATA_TRAIN_Filtered$GameWeather))
# sum(NFL_DATA_TRAIN_Filtered$GameWeather == "")

# Factor the DOWNS, Ordinally
NFL_DATA_TRAIN_Filtered$Down = factor(as.numeric(NFL_DATA_TRAIN_Filtered$Down), order = TRUE, levels = c(1,2,3,4,5))

# Player WEIGHT 
NFL_DATA_TRAIN_Filtered$PlayerWeight = as.numeric(NFL_DATA_TRAIN_Filtered$PlayerWeight) 

# Factor POSITION
NFL_DATA_TRAIN_Filtered$Position = factor(NFL_DATA_TRAIN_Filtered$Position)

# factor POSITION
NFL_DATA_TRAIN_Filtered$Position = factor(NFL_DATA_TRAIN_Filtered$Position)

# factor SEASON
NFL_DATA_TRAIN_Filtered$Season = factor(NFL_DATA_TRAIN_Filtered$Season)


# DATA CLEANING (REMOVING NA'S, and observations that happen less than 3 times)
# This was causing issues where one fold has a factor but another fold does not

# Need to delete a row within a column if there is just 1 special case. (Minimun 3 observations)

# DefensePersonnel (reduces observations by 11)
NFL_DATA_TRAIN_Filtered = NFL_DATA_TRAIN_Filtered[unsplit(table(NFL_DATA_TRAIN_Filtered$DefensePersonnel), NFL_DATA_TRAIN_Filtered$DefensePersonnel) >= 3, ]

# Same for OffensePersonnel (reduces by 18 observations)
NFL_DATA_TRAIN_Filtered = NFL_DATA_TRAIN_Filtered[unsplit(table(NFL_DATA_TRAIN_Filtered$OffensePersonnel), NFL_DATA_TRAIN_Filtered$OffensePersonnel) >= 3, ]

# Same for Position
NFL_DATA_TRAIN_Filtered = NFL_DATA_TRAIN_Filtered[unsplit(table(NFL_DATA_TRAIN_Filtered$Position), NFL_DATA_TRAIN_Filtered$Position) >= 3, ]

# Same for Defenders In the Box 
NFL_DATA_TRAIN_Filtered = NFL_DATA_TRAIN_Filtered[unsplit(table(NFL_DATA_TRAIN_Filtered$DefendersInTheBox), NFL_DATA_TRAIN_Filtered$DefendersInTheBox) >= 3, ]

# Same for Offense Formation
NFL_DATA_TRAIN_Filtered = NFL_DATA_TRAIN_Filtered[unsplit(table(NFL_DATA_TRAIN_Filtered$OffenseFormation), NFL_DATA_TRAIN_Filtered$OffenseFormation) >= 3, ]

# Need to remove NA Rows: Still 16,758 observations out of ~18,000
NFL_DATA_TRAIN_Filtered_Final <- na.omit(NFL_DATA_TRAIN_Filtered)
# View(NFL_DATA_TRAIN_Filtered_Final)
```



## BEGINNING OF ANALYSIS

# Regression with all Covariates
```{r}
NFL_Train_Total_Model = lm(Yards ~ ., data=NFL_DATA_TRAIN_Filtered_Final)
NFL_Train_Total_Model.cv = cvFit(NFL_Train_Total_Model, data=NFL_DATA_TRAIN_Filtered_Final, y=NFL_DATA_TRAIN_Filtered_Final$Yards, K=10, seed = 123)
NFL_Train_Total_Model.cv # RMSE= 6.388  # May have collinearity
NFL_Train_Total_Model$coefficients
```

```{r}
#NFL_Train_Total_Model$coef)
#summary(NFL_Train_Total_Model)$coef[summary(NFL_Train_Total_Model)$coef[,4] <= .05, 1]
data.frame(summary(NFL_Train_Total_Model)$coef)
```


# What if we always predicted the mean of yards? (Just Intercept Term)
``` {r}
NFL_Train_Total_Model1 = lm(Yards ~ 1, data=NFL_DATA_TRAIN_Filtered_Final)
NFL_Train_Total_Model1.cv = cvFit(NFL_Train_Total_Model1, data=NFL_DATA_TRAIN_Filtered_Final, y=NFL_DATA_TRAIN_Filtered_Final$Yards, K=10, seed = 123)
NFL_Train_Total_Model1.cv # RMSE= 6.4191 
``` 

# Forward Stepwise Regression

```{r}
min_model = NFL_Train_Total_Model1
max_model = NFL_Train_Total_Model
stepwise_model = step(min_model, direction='forward', scope=max_model)
summary(stepwise_model)
# Ultimately, this is saying that the extra info we gain is not worth the complexity
```


# Backward Stepwise Regerssion
``` {r echo = T, results = 'hide'}
backward_step = step(max_model, direction='backward')
backward_step 
```

``` {r}
backward_step.cv = cvFit(backward_step, data=NFL_DATA_TRAIN_Filtered_Final, y=NFL_DATA_TRAIN_Filtered_Final$Yards, K=10, seed = 123)
backward_step.cv # RMSE= 6.5386 # Best Model so far
```

# Ridge Regression
```{r}
library(glmnet)
## NORMALIZE Continouous Covariates # Will have to normalize defenders in box if we make continuous
Standardized_NFL_TRAIN = NFL_DATA_TRAIN_Filtered_Final
Standardized_NFL_TRAIN$X = scale(Standardized_NFL_TRAIN$X)
Standardized_NFL_TRAIN$Y = scale(Standardized_NFL_TRAIN$Y)
Standardized_NFL_TRAIN$GameClock = scale(Standardized_NFL_TRAIN$GameClock )
Standardized_NFL_TRAIN$Distance = scale(Standardized_NFL_TRAIN$Distance)
Standardized_NFL_TRAIN$PlayerHeight = scale(Standardized_NFL_TRAIN$PlayerHeight)
Standardized_NFL_TRAIN$PlayerWeight = scale(Standardized_NFL_TRAIN$PlayerWeight)
Standardized_NFL_TRAIN$Week = scale(Standardized_NFL_TRAIN$Week)
Standardized_NFL_TRAIN$Temperature = scale(Standardized_NFL_TRAIN$Temperature)
Standardized_NFL_TRAIN$Humidity = scale(Standardized_NFL_TRAIN$Humidity)
Standardized_NFL_TRAIN$TimeDifference = scale(Standardized_NFL_TRAIN$TimeDifference)
Standardized_NFL_TRAIN$HomeScoreAdvantage = scale(Standardized_NFL_TRAIN$HomeScoreAdvantage )
Standardized_NFL_TRAIN$PlayerAge = scale(Standardized_NFL_TRAIN$PlayerAge)

# Ridge Regression 
# Ridge alpha = 0
x = model.matrix(Yards~. , Standardized_NFL_TRAIN)
y = Standardized_NFL_TRAIN$Yards
ridge_mod = glmnet(x, y, alpha = 0)
# install.packages("plotmo")
plot_glmnet(ridge_mod, label = TRUE) 
```

``` {r}
cvfit = cv.glmnet(x, y, alpha = 0)
plot(cvfit) 
bestlam = cvfit$lambda.min # = 4.9384
```

# Get coefficients when log lamda is 10.098

```{r}
y_predicted <- predict(cvfit, s = bestlam, newx = x) # same x, in sample prediction
ridge_RMSE = sqrt(mean((y_predicted - y)^2))
# ridge_RMSE # = 6.3333
#coef(ridge_mod)[,4.9384] # Best is again basically forcing all the betas to 0. Just predict mean
```

# Lasso Regression

```{r}
lasso_mod= glmnet(x, y, alpha = 1)
# coef(lasso_mod)[,50]
plot_glmnet(lasso_mod, label = TRUE)
```

``` {r}
cvfit_lasso = cv.glmnet(x, y, alpha = 1)
plot(cvfit_lasso) 
bestlam_lasso = cvfit_lasso$lambda.min # = 0.046056
```

``` {r}
y_predicted_lasso <- predict(cvfit_lasso, s = bestlam_lasso, newx = x) # same x, in sample prediction
lasso_RMSE = sqrt(mean((y_predicted_lasso - y)^2))
lasso_RMSE # = 6.3349
```

``` {r}
out = glmnet(x, y, alpha = 1) # Fit ridge regression model on full dataset
#predict(out, type = "coefficients", s = bestlam_lasso)[1:80,] # Display coefficients using lambda chosen by CV. Forces everything to 0
```

































#Applying same transformation to test set

```{r}

# Function to take the difference in time from the dataframe
timeDifference <- function(time) {
  num <- gsub("[:]", "" , str_sub(time, 12, 19), perl=TRUE)
  hr <- ifelse(str_sub(num, 1, 2) == "00", 24, as.numeric(str_sub(num, 1, 2)))
  min <- as.numeric(str_sub(num, 3, 4))
  sec <- as.numeric(str_sub(num, 5, 6))
  newTime <- 3600*hr + 60 * min + sec
  return(newTime)
}
# Add Time_Difference between the snap and the handoff
NFL_DATA_Holdout$TimeDifference <- 
  timeDifference(NFL_DATA_Holdout$TimeHandoff) - timeDifference(NFL_DATA_Holdout$TimeSnap)

# Add the Difference in Score by home score - visitor score
# Difference in Score (Pair with which team is winning (HomeScore-AwayScore))
NFL_DATA_Holdout$HomeScoreAdvantage <- 
  NFL_DATA_Holdout$HomeScoreBeforePlay - NFL_DATA_Holdout$VisitorScoreBeforePlay

# Add the age of the running player

# Change the birth dates to strings
NFL_DATA_Holdout$PlayerBirthDate = as.character(NFL_DATA_Holdout$PlayerBirthDate)
# Grab the Year for each of the running player
Birth_Year = str_sub(NFL_DATA_Holdout$PlayerBirthDate, 7, 11)
# Grab Month of each running player
Birth_Month = str_sub(NFL_DATA_Holdout$PlayerBirthDate, 1, 2)
# If Born in July (07) Have lived 5/12 of a year. ie (12 - (Birth_Month)) / 12
How_Much_Of_Year_Lived = (12 - as.numeric(Birth_Month)) / 12
Years_Lived = NFL_DATA_Holdout$Season - as.numeric(Birth_Year)
Total_Years_Lived = Years_Lived + How_Much_Of_Year_Lived
NFL_DATA_Holdout$PlayerAge = Total_Years_Lived

# Change HEIGHT to inches and continuous
Feet = as.numeric(str_sub(NFL_DATA_Holdout$PlayerHeight, 1, 1)) 
Inches = as.numeric(str_sub(NFL_DATA_Holdout$PlayerHeight, 3, 4)) 
Heights = (Feet * 12) + Inches
NFL_DATA_Holdout$PlayerHeight = Heights

# Changes GAMECLOCK to Seconds. 

NFL_DATA_Holdout$GameClock = as.numeric(NFL_DATA_Holdout$GameClock)

# FACTORING VARIABLES INTO CATEGORICAL

# Factor OFFENSE FORMATION 
NFL_DATA_Holdout$OffenseFormation = factor(NFL_DATA_Holdout$OffenseFormation) 

# DEFENDERS IN BOX (Need Categorical and Ordinal)
NFL_DATA_Holdout$DefendersInTheBox = factor(NFL_DATA_Holdout$DefendersInTheBox)  


# REFACTOR DEFENDER IN BOX TO INCLUDE ORDINALITY
# Leaving them unordered was for graphs above, to look good
NFL_DATA_Holdout$DefendersInTheBox = factor(NFL_DATA_Holdout$DefendersInTheBox, order = TRUE, levels= c(2,3,4,5,6,7,8,9,10,11)) 

# Drop columns that are collinear, or we think are not critical to our model
NFL_DATA_HOLDOUT = select(NFL_DATA_Holdout, 
                                -GameId, -PlayId, -Team, -S, -A, -Dis,
                                -Orientation, -Dir, -DisplayName, -JerseyNumber,
                                -YardLine, -FieldPosition, -HomeScoreBeforePlay,
                                -VisitorScoreBeforePlay, -NflId, -TimeHandoff,
                                -TimeSnap, -PlayerBirthDate, -PlayerCollegeName, 
                                -Location,-WindSpeed, -WindDirection, -StadiumType,
                                -Turf,-GameWeather,-NflIdRusher, -Stadium) 
# Turf and stadium type all captured in stadium
# View(NFL_DATA_TRAIN_Filtered) # drop game weather, captured in Week, and Stadium. Also too many missing values



# Need to count how many NA / Empty cells there are for each column
# summary(NFL_DATA_TRAIN_Filtered)  # changed empty to NA when reading in file
# GameWeather, Temperature, Humidy all have missing or NA data
# sum(is.na(NFL_DATA_TRAIN_Filtered$GameWeather))
# sum(NFL_DATA_TRAIN_Filtered$GameWeather == "")

# Factor the DOWNS, Ordinally
NFL_DATA_HOLDOUT$Down = factor(as.numeric(NFL_DATA_HOLDOUT$Down), order = TRUE, levels = c(1,2,3,4,5))

# Player WEIGHT 
NFL_DATA_HOLDOUT$PlayerWeight = as.numeric(NFL_DATA_HOLDOUT$PlayerWeight) 

# Factor POSITION
NFL_DATA_HOLDOUT$Position = factor(NFL_DATA_HOLDOUT$Position)

# factor POSITION
NFL_DATA_HOLDOUT$Position = factor(NFL_DATA_HOLDOUT$Position)

# factor SEASON
NFL_DATA_HOLDOUT$Season = factor(NFL_DATA_HOLDOUT$Season)


# DATA CLEANING (REMOVING NA'S, and observations that happen less than 3 times)
# This was causing issues where one fold has a factor but another fold does not

# Need to delete a row within a column if there is just 1 special case. (Minimun 3 observations)

# DefensePersonnel (reduces observations by 11)
NFL_DATA_HOLDOUT = NFL_DATA_HOLDOUT[unsplit(table(NFL_DATA_HOLDOUT$DefensePersonnel), NFL_DATA_HOLDOUT$DefensePersonnel) >= 3, ]

# Same for OffensePersonnel (reduces by 18 observations)
NFL_DATA_HOLDOUT = NFL_DATA_HOLDOUT[unsplit(table(NFL_DATA_HOLDOUT$OffensePersonnel), NFL_DATA_HOLDOUT$OffensePersonnel) >= 3, ]

# Same for Position
NFL_DATA_HOLDOUT = NFL_DATA_HOLDOUT[unsplit(table(NFL_DATA_HOLDOUT$Position), NFL_DATA_HOLDOUT$Position) >= 3, ]

# Same for Defenders In the Box 
NFL_DATA_HOLDOUT = NFL_DATA_HOLDOUT[unsplit(table(NFL_DATA_HOLDOUT$DefendersInTheBox), NFL_DATA_HOLDOUT$DefendersInTheBox) >= 3, ]

# Same for Offense Formation
NFL_DATA_HOLDOUT = NFL_DATA_HOLDOUT[unsplit(table(NFL_DATA_HOLDOUT$OffenseFormation), NFL_DATA_HOLDOUT$OffenseFormation) >= 3, ]

# Need to remove NA Rows: Still 16,758 observations out of ~18,000
NFL_DATA_HOLDOUT_Final <- na.omit(NFL_DATA_HOLDOUT)
# View(NFL_DATA_TRAIN_Filtered_Final)
```





# Using model with all covariates to predict on test set 
```{r}
NFL_TEST_Model.cv = cvFit(NFL_Train_Total_Model, data=NFL_DATA_HOLDOUT_Final, y=NFL_DATA_HOLDOUT_Final$Yards, K=10, seed = 122)
NFL_TEST_Model.cv # RMSE= 6.4191 

#
#pred = predict(NFL_Train_Total_Model, NFL_DATA_HOLDOUT_Final)
#rmse.test = sqrt(mean((pred - NFL_DATA_HOLDOUT_Final$Yards)^2))
#rmse.test
```



# Using model with only intercept term
```{r}
NFL_TEST_Model.cv = cvFit(NFL_Train_Total_Model1, data=NFL_DATA_HOLDOUT_Final, y=NFL_DATA_HOLDOUT_Final$Yards, K=10, seed = 122)
NFL_TEST_Model.cv # RMSE= 6.4191 
```


#Using 'best' model (backward stepwise regression)
```{r}
NFL_TEST_Model.cv = cvFit(backward_step, data=NFL_DATA_HOLDOUT_Final, y=NFL_DATA_HOLDOUT_Final$Yards, K=10, seed = 122)
NFL_TEST_Model.cv # RMSE= 6.4191 
```

```{r}
summary(backward_step)
```

# Fitting Backward Stepwise Regression to test set

``` {r}
#NFL_TEST_min_Model = lm(Yards ~ 1, data=NFL_DATA_HOLDOUT_Final)
NFL_TEST_max_Model = lm(Yards ~ ., data=NFL_DATA_HOLDOUT_Final)
test_backward_step = step(NFL_TEST_max_Model, direction='backward')
```

```{r}
cat("Stepwise Backward Model on test set")
summary(test_backward_step)
```




```{r}
# Bootsrap sample distribution holders

Season2018_Hold = rep(0, 1000) # 2
Distance_Hold = rep(0, 1000) # 3
Defender_L = rep(0, 1000)
Defender_Q = rep(0, 1000)
Defender_C = rep(0, 1000)
Df_4 = rep(0, 1000)
Df_5 = rep(0, 1000)
Df_6 = rep(0, 1000)
Df_7 = rep(0, 1000)
Df_8 = rep(0, 1000)
FB = rep(0, 1000)
HB = rep(0, 1000)
QB = rep(0, 1000)
RB = rep(0, 1000)
TE = rep(0, 1000)
WR = rep(0, 1000)
Temp_Hold = rep(0, 1000)
Humidty_Hold = rep(0, 1000)
Time_Diff_Hold = rep(0, 1000)
Home_Adva_Hold = rep(0, 1000)
Age_Hold = rep(0, 1000)

```


# Bootstrap 
```{r}
# backward_step  # This is the BackwardStep Model
# NFL_DATA_TRAIN_Filtered_Final # this is where the data is kept
# NFL_Train_Total_Model = lm(Yards ~ ., data=NFL_DATA_TRAIN_Filtered_Final)
# summary(model_boot)$coefficients[2]

set.seed(123)
for(i in 1:1000){
  Sample_Train_Data = sample_n(NFL_DATA_TRAIN_Filtered_Final, 16761, replace=TRUE)
  model_boot = lm(Yards ~ Season + Distance + DefendersInTheBox + 
                  Position + Temperature + Humidity + TimeDifference + 
                  HomeScoreAdvantage + PlayerAge, data=Sample_Train_Data)
  # Grab the coefficient
  Season2018_Hold[i] = summary(model_boot)$coefficients[2]
  Distance_Hold[i] = summary(model_boot)$coefficients[3]#3
  Defender_L[i] = summary(model_boot)$coefficients[4]
  Defender_Q[i] = summary(model_boot)$coefficients[5]
  Defender_C[i] = summary(model_boot)$coefficients[6]
  Df_4[i] = summary(model_boot)$coefficients[7]
  Df_5[i] = summary(model_boot)$coefficients[8]
  Df_6[i] = summary(model_boot)$coefficients[9]
  Df_7[i] = summary(model_boot)$coefficients[10]
  Df_8[i] = summary(model_boot)$coefficients[11]
  FB[i] = summary(model_boot)$coefficients[12]
  HB[i] = summary(model_boot)$coefficients[13]
  QB[i] = summary(model_boot)$coefficients[14]
  RB[i] = summary(model_boot)$coefficients[15]
  TE[i] = summary(model_boot)$coefficients[16]
  WR[i] = summary(model_boot)$coefficients[17]
  Temp_Hold[i] = summary(model_boot)$coefficients[18]
  Humidty_Hold[i] = summary(model_boot)$coefficients[19]
  Time_Diff_Hold[i] = summary(model_boot)$coefficients[20]
  Home_Adva_Hold[i] = summary(model_boot)$coefficients[21]
  Age_Hold[i] = summary(model_boot)$coefficients[22]
}

```

```{r}

calculate_CI <- function(dist){
  se_bootrap = sd(dist) / sqrt(length(dist))
  min_conf = mean(dist) - (1.96 * se_bootrap)
  max_conf = mean(dist) + (1.96 * se_bootrap)
  # Quantile Bootstrap
  sorted_dist = sort(dist)
  min = sorted_dist[.025 * length(sorted_dist)]
  max = sorted_dist[.975 * length(sorted_dist)]
  return(c(min_conf,max_conf, min,max))
}

```

```{r}
calculate_CI(Season2018_Hold)
calculate_CI(Distance_Hold)
calculate_CI(Defender_L)
calculate_CI(Defender_Q)
calculate_CI(Defender_C)DT
calculate_CI(Df_4)
calculate_CI(Df_5)
calculate_CI(Df_6)
calculate_CI(Df_7)
calculate_CI(Df_8)
calculate_CI(FB)
calculate_CI(HB)
calculate_CI(QB)
calculate_CI(RB)
calculate_CI(TE)
calculate_CI(WR)
calculate_CI(Temp_Hold)
calculate_CI(Humidty_Hold)
calculate_CI(Time_Diff_Hold)
calculate_CI(Home_Adva_Hold)
calculate_CI(Age_Hold)
```
```{r}
confint(backward_step)
```
 
```{r}

ggplot() + 
  geom_histogram(mapping = aes(x = Defender_L), bins = 50) + 
  ggtitle("Distribution of Defender_L")

ggplot() + 
  geom_histogram(mapping = aes(x = Home_Adva_Hold), bins = 50) + 
  ggtitle("Distribution of Home_Score_Advantage")

```





















# Classification. Outcome Variable: NFL Yards >= distance. (Whether or not they get a first down)

# With all Covariates
```{r}
NFL_Train_Total_Model_c <- NFL_DATA_TRAIN_Filtered_Final
NFL_Train_Total_Model_c$FirstDown <- ifelse(
  NFL_Train_Total_Model_c$Yards >= NFL_Train_Total_Model_c$Distance, 1, 0)
NFL_Train_Total_Model_c$FirstDown <-
  factor(NFL_Train_Total_Model_c$FirstDown)
NFL_Train_Total_Model_c <- select(NFL_Train_Total_Model_c, -Yards) #remove colinear response variable from regression
```

# Implementing Cross-Fold Validation for Classification
```{r}
set.seed(122)
f <- createFolds(y=NFL_Train_Total_Model_c$FirstDown, k=10)
train_fold <- function (i) {
  NFL_Train_Total_Model_c[-unlist(f[i]),]
}

test_fold <- function (i) {
  NFL_Train_Total_Model_c[unlist(f[i]),]
}

```

# Baseline model
```{r}
accuracyR <- c()

for (i in 1:10) {
  glm_Model = glm(FirstDown ~ ., data=train_fold(i), 
                              family = binomial)
  predict_result <- predict(glm_Model, newdata=test_fold(i), type="response")
  predict_logit <- ifelse(predict_result > 0.5, 1, 0)
  t <- table(predict_logit, test_fold(i)$FirstDown)
  accuracyR[i] = (t[1,1]+t[2,2])/dim(test_fold(i))[1]
}
cat("Mean accuracy of 10-fold validation is: ", mean(accuracyR), "\n")
confusionMatrix(test_fold(i)$FirstDown, factor(predict_logit))
```

```{r eval=FALSE}
  #Why you don't use in-prediction error - biased -lower error or higher accuracy
  glm_Model = glm(FirstDown ~ ., data=NFL_Train_Total_Model_c, 
                              family = binomial)
  predict_result <- predict(glm_Model, newdata=NFL_Train_Total_Model_c, type="response")
  predict_logit <- ifelse(predict_result > 0.5, 1, 0)
  t <- table(predict_logit, NFL_Train_Total_Model_c$FirstDown)
  accuracyR = (t[1,1]+t[2,2])/dim(NFL_Train_Total_Model_c)[1] #0.82438
```



```{r, include=FALSE, eval=FALSE}
#Penalized Logistic Regression
#Ridge
#error in prediction
#reuse standardized covarites from lasso in regresion

standardized_new <- Standardized_NFL_TRAIN
standardized_new$FirstDown <- NFL_Train_Total_Model_c$FirstDown
standardized_new <- select(standardized_new, -Yards) #remove colinear variable 

f <- createFolds(y=standardized_new$FirstDown, k=10)
s_train_fold <- function (i) {
  NFL_Train_Total_Model_c[-unlist(f[i]),]
}

s_test_fold <- function (i) {
  NFL_Train_Total_Model_c[unlist(f[i]),]
}
```

```{r, include=FALSE, eval=FALSE}
accuracyR <- c()
for (i in 1:10) {
  x = model.matrix(FirstDown~., s_train_fold(i))
  y = s_train_fold(i)$FirstDown
  cvfit_ridge = cv.glmnet(x, y, alpha = 0, family="binomial")
  #plot(cvfit_ridge) 
  lam = cvfit_ridge$lambda.min #
  model <- glmnet(x, y, alpha = 0, family = "binomial",
                lambda = lam)
  x_pred = model.matrix(FirstDown~., s_test_fold(i))
  y_predicted <- model %>% predict(newx = x_pred) # same x, in sample prediction
  predict_logit <- ifelse(y_predicted > 0.5, 1, 0)
  t <- table(predict_logit, s_test_fold(i)$FirstDown)
  #accuracyR <- (t[1,1]+t[2,2])/dim(standardized_new)[1]  
  accuracyR[i] = (t[1,1]+t[2,2])/dim(s_test_fold(i))[1]
}
cat("Mean accuracy of 10-fold validation is: ", mean(accuracyR))
#print(mean(accuracyR)) #0.81483
confusionMatrix(test_fold(i)$FirstDown, factor(predict_result))

```



# Using classification model to calculate test error

# With all Covariates
```{r}
NFL_HOLDOUT_C <- NFL_DATA_HOLDOUT_Final
NFL_HOLDOUT_C$FirstDown <- ifelse(
  NFL_HOLDOUT_C$Yards >= NFL_HOLDOUT_C$Distance, 1, 0)
NFL_HOLDOUT_C$FirstDown <-
  factor(NFL_HOLDOUT_C$FirstDown)
NFL_HOLDOUT_C <- select(NFL_HOLDOUT_C, -Yards) #remove colinear response variable from regression
```

# Implementing Cross-Fold Validation for Classification
```{r}
set.seed(121)
f <- createFolds(y=NFL_HOLDOUT_C$FirstDown, k=10)
train_fold <- function (i) {
  NFL_HOLDOUT_C[-unlist(f[i]),]
}

test_fold <- function (i) {
  NFL_HOLDOUT_C[unlist(f[i]),]
}

```

# Baseline model
```{r}
accuracyR <- c()

for (i in 1:10) {
  glm_Model = glm(FirstDown ~ ., data=train_fold(i), 
                              family = binomial)
  predict_result <- predict(glm_Model, newdata=NFL_HOLDOUT_C, type="response")
  predict_logit <- ifelse(predict_result > 0.5, 1, 0)
  t <- table(predict_logit, NFL_HOLDOUT_C$FirstDown)
  accuracyR[i] = (t[1,1]+t[2,2])/dim(NFL_HOLDOUT_C)[1]
}
cat("Mean accuracy of 10-fold validation is: ", mean(accuracyR), "\n")

#confusionMatrix(NFL_HOLDOUT_C$FirstDown, factor(predict_logit))
```


